---
title: 'Implementing solutions: A review exploring determinants for the implementation of solutions to food-energy-water system challenges'
subtitle: 'Supplemental materials for submittal to Frontiers in Environmental Science'
author: 
  - Paula Williams
  - Jamie Trammell
  - Leah Jones-Crank
  - Ruchie Pathak
  - Erich Seamon
  - Alyssa Thomas
  - Dan Cronan
  - Andrew Kliskey
date: "`r Sys.Date()`"
always_allow_html: yes
output:
  pdf_document:
    toc: yes
  includes:
      in_header: header.tex
params: null
mainfont: serif
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 120)
```

\newpage

# Supplemental Materials Summary

This analysis focuses on examining if how stakeholder engagement, and the level of engagement, impacts whether a solution for research outcomes is proposed and/or implemented.  This meta-synthesis of 483 papers were evaluated and coded using several differing engagement scales. Additionally, each paper was coded by the geographic scale, and whether a computational model was used as part of the research. 



```{r echo=FALSE, warning=FALSE, message=FALSE}

library(ggplot2)
library(epiDisplay)
library(caret)

#crcdata <- read.csv("./data/crc_data_ES2.csv")
crcdata <- read.csv("./data/crc_data_aug25.csv")

#remove citation

crcdata <- crcdata[, -which(names(crcdata) == "Citation")]

#set YN to 01
require(dplyr)
crcdata <- crcdata %>%
  mutate(solution_proposed_YN = ifelse(solution_proposed_YN == "N",0,1))

require(dplyr)
crcdata <- crcdata %>%
  mutate(solution_implemented_YN = ifelse(solution_implemented_YN == "N",0,1))

require(dplyr)
crcdata <- crcdata %>%
  mutate(S_stakeholder_engagement_YN = ifelse(S_stakeholder_engagement_YN == "N",0,1))

#fix colnames
#names(crcdata)[names(crcdata) == 'STE_IAP2_NA'] <- 'STE_IAP2_data_gathering'
#names(crcdata)[names(crcdata) == 'G_na'] <- 'G_nogeography'
names(crcdata)[names(crcdata) == 'Year'] <- 'year'
#names(crcdata)[43] <- "SC_researcher"

crcdata <- head(crcdata,-1)

#names(crcdata)[51] <- "G_notdescribed"


```


\newpage

# Variable Summary

Below is a list of the categorical variables generated from the literature reviews

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
df <- data.frame(Variables = c("Year", "Solution Proposed", "Solution Implemented", "Solution Type", "Researcher Type", "Stakeholder Type", "Stakeholder engagement Scale - Ghodsvali", "Geographical Type", "Region"), Description = linebreak(c("Year of citation", "Was a solution proposed?", "Was a solution implemented?","If a solution was proposed, what was the solution type? Groups include: Technology, Policy, Institutional, Social, Economic, Ecological, and Educational.", "What was the research type?  Groups include: NGO, English, Math, Computer Science, Physics, Engineering, Interdisciplinary, Social Science, Economics, Agriculture, and Other","What was the stakeholder type? Groups include: Farmers, Combined Government, Combined Coalition, Combined Industry, Migrants, Youth, Public, Univerity, and Experts", "If a stakeholder was engaged, categorization of the engagement using the Ghodsvali scale. Groups include: Nominal, Instrumental, Representation, and Transformative", "What the geography type? Groups include: Not Described, Local, Regional, National, Multinational, Global, and No Geography", "What was the country?")))
kable(df, col.names = c("Variable Name", "Description"), escape = F, caption = "Table T1: Variable Descriptions") %>% column_spec(2, width = "2in") %>% column_spec(c(1:2), width = "2.5in") %>% kable_styling(full_width = FALSE,latex_options = c("hold_position", "scale_down"), font_size = 8) %>% collapse_rows(columns = 2) %>% kable_classic_2(full_width = F) %>% row_spec(0,bold=TRUE)

```


<!-- \newpage -->

<!-- # Variable Summary REFINED -->

<!-- Below is a list of the categorical variables generated from the literature reviews REFINED -->

<!-- ```{r echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- library(knitr) -->
<!-- library(kableExtra) -->
<!-- df <- data.frame(Variables = c("Year", "Solution Proposed", "Solution Implemented",  "Stakeholder Type", "Stakeholder engagement Scale - Ghodsvali", "Stakeholder engagement Scale - IAP2", "Stakeholder Engagement Scale - Local"), Description = linebreak(c("Year of citation", "Was a solution proposed?", "Was a solution implemented?","What was the stakeholder type? Groups include: Farmers, Combined Government, Combined Coalition, Combined Industry, Migrants, Youth, Public, Univerity, and Experts", "If a stakeholder was engaged, categorization of the engagement using the Ghodsvali scale. Groups include: Nominal, Instrumental, Representation, and Transformative", "If a stakeholder was engaged, categorization of the engagement using the IAP2 scale. Groups include: Data Gathering, Inform, Consult, Involve, Collaborate, and Empower", "If a stakeholder was engaged, categorization of the engagement using a customized scale. Groups include: Researcher, Data Gathering, Inform, Perspectives, Planning, Identify, Envision, and Implement"))) -->
<!-- kable(df, col.names = c("Variable Name", "Description"), escape = F, caption = "Table T1: Variable Descriptions") %>% column_spec(2, width = "2in") %>% column_spec(c(1:2), width = "2.5in") %>% kable_styling(full_width = FALSE,latex_options = c("hold_position", "scale_down"), font_size = 8) %>% collapse_rows(columns = 2) %>% kable_classic_2(full_width = F) %>% row_spec(0,bold=TRUE)  -->

<!-- ``` -->


\newpage

# Chi-Square Testing
## Chi Square Testing: solution proposed or not vs. stakeholder engagement
3
Chi Square and Fishers Exact Test on contingency table with Solution/No Solution as the explanatory variable, and engaged stakeholder/did not engage stakeholder as the response variable. 

**ChiSquare = 46: Fishers Exact Test Odds Ratio: 17: Not Independent**

Both chi square and fishers exact test were significant, with a chi square approximation of ~43, which is well above the critical value (3.84 with one degree of freedom).  Fishers Exact Test returned an odds ratio of ~17. The alternative hypothesis: true odds ratio is not equal to 1, therefore the null hypothesis is rejected - the groups are not independent. 

The Fishers Exact Test defaults to associating the odds ratio (which can represent effect size) with the first cell. In this instance "The odds of having a solution is 17 times that for an engaged stakeholder".  You could flip the response and explanatory variables, but the odds ratio would stay the same.

For more info on this topic see: Kim HY. Statistical notes for clinical researchers: Chi-squared test and Fisher's exact test. Restor Dent Endod. 2017 May;42(2):152-155. doi: 10.5395/rde.2017.42.2.152. Epub 2017 Mar 30. PMID: 28503482; PMCID: PMC5426219.


```{r, echo=FALSE}
library(Barnard)
library(vcd)


crcdata$solution_proposed_YN <- as.factor(crcdata$solution_proposed_YN)
crcdata$S_stakeholder_engagement_YN <- as.factor(crcdata$S_stakeholder_engagement_YN)
crcdata$solution_implemented_YN <- as.factor(crcdata$solution_implemented_YN)

levels(crcdata$S_stakeholder_engagement_YN) = c("N", "Y")
levels(crcdata$S_stakeholder_engagement_YN) = c("N", "Y")
levels(crcdata$solution_proposed_YN) = c("N", "Y")
levels(crcdata$solution_implemented_YN) = c("N", "Y")

crcdata_table <- table(crcdata$solution_proposed_YN, crcdata$S_stakeholder_engagement_YN)
#colnames(crcdata_table) <- c("NE", "E")

solution_stakeholder <- crcdata_table

names(dimnames(solution_stakeholder)) <- c('solution','stakeholder')
solution_stakeholder

solution_stakeholder <- as.table(solution_stakeholder)

summary(solution_stakeholder)

fisher.test(solution_stakeholder)
barnard.test(76, 465, 14, 18)
assocstats(solution_stakeholder)

```

\newpage

# Summary Statistics Graphs

## Were solutions proposed in the set of all papers?

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(crcdata,aes(x=solution_proposed_YN, fill=year)) + ggtitle("Solution Proposed: Y=18") + labs(x="Was a solution proposed?") +
  geom_bar(stat = "count", position = "dodge", fill="#99CC99") + theme_minimal() + theme_minimal()+
  theme(text = element_text(),
        axis.text.x = element_text(angle = 0, hjust = 1)) 

```
\newpage

## Were solutions implemented in the set of all papers?

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(crcdata,aes(x=solution_implemented_YN, fill=year)) + ggtitle("Solution Implemented: Y=11") + labs(x="Was a solution implemented?") + geom_bar(stat = "count", position = "dodge", fill="#99CC99") + theme_minimal()+
  theme(text = element_text(),
        axis.text.x = element_text(angle = 0, hjust = 1)) 

```
\newpage

## What were the solution types?

```{r echo=FALSE, warning=FALSE, message=FALSE}

# par(mar = c(8,4,4,4))
# plot3 <- barplot(colSums(crcdata[,4:10]), las=2, main="Solution Types: N=18", names.arg = c("Technological", "Policy", "Institutional", "Social", "Economic", "Ecological", "Educational"))

plot3_data <- colSums(crcdata[,4:10])
p3a <- as.data.frame(plot3_data)
p3a$group <- c("Technological", "Policy", "Institutional", "Social", "Economic", "Ecological", "Educational")

# 
ggplot(p3a, aes(x=group, y=plot3_data)) + geom_bar(stat="identity",fill="#99CC99") + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  xlab("Solution Types") + ylab("Count")


```
\newpage

## Were stakeholders engaged?

```{r echo=FALSE, warning=FALSE, message=FALSE}

# nrow(subset(crcdata, solution_proposed_YN == "1"))/nrow(subset(crcdata, solution_proposed_YN == "0"))
# 
# nrow(subset(crcdata, solution_implemented_YN == "1"))/nrow(subset(crcdata, solution_implemented_YN == "0"))


#stakeholder
#par(mar = c(4,4,4,4))
#barplot(table(crcdata$S_stakeholder_engagement_YN), main = "Stakeholder engagement: Y=92")

ggplot(crcdata,aes(x=S_stakeholder_engagement_YN, fill=year)) + ggtitle("Stakeholder Engagement: YES=92") + labs(x="Were stakeholders engaged?") +
  geom_bar(stat = "count", position = "dodge", fill="#99CC99") + theme_minimal()

```
\newpage

## All FEWS papers by year

```{r echo=FALSE, warning=FALSE, message=FALSE}
#year
crcdata_remove2011 <- subset(crcdata, year != "2011")
ggplot(crcdata_remove2011, aes(x=year)) + 
    geom_bar(stat="count")+theme_bw() + scale_x_continuous(breaks = crcdata$year) +
  geom_bar(stat = "count", position = "dodge", fill="#99CC99") + theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```


```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
#year

png("./images/fews_year.png", res=300, width = 1500, height = 1500 )
ggplot(crcdata_remove2011, aes(x=year)) + 
    geom_bar(stat="count")+theme_bw() + scale_x_continuous(breaks = crcdata$year) +
  geom_bar(stat = "count", position = "dodge", fill="#99CC99") + theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
dev.off()


```


\newpage

## Level of stakeholder engagement by year - Ghodsvali scale

```{r echo=FALSE, warning=FALSE, message=FALSE}
#year
library(dplyr)

crcdata_new <- crcdata %>% 
  mutate(ghodsvali = sprintf("%d_%d_%d_%d", STE_G_nominal, STE_G_instrumental, STE_G_representation, STE_G_transformative))

crcdata_new <- crcdata_new[!(crcdata_new$ghodsvali %in% "0_0_0_0"),]

library(dplyr)
crcdata_new <- crcdata_new %>%
    mutate(ghodsvali = recode(ghodsvali, '0_0_0_0' = 'none', '1_0_0_0' = 'nominal', '0_1_0_0' =  'instrumental', '0_0_1_0' = 'representative',  '0_0_0_1' = 'transformative' ))

crcdata_new$ghodsvali <- as.factor(crcdata_new$ghodsvali)


crcdata_new$ghodsvali <- factor(crcdata_new$ghodsvali, levels = c("nominal", "instrumental", "representative", "transformative"))

my_colors <- c("#99CC99", "#669999", "#6666CC", "#CC9966")

ggplot(crcdata_new, aes(fill=ghodsvali, x=year)) + 
    geom_bar(position="stack", stat="count")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + scale_x_continuous(breaks = crcdata_new$year) + scale_fill_manual(values = my_colors, name = "Ghodsvali Scale")


```


```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}

png("./images/fews_ghodsvali.png", res=300, width = 1500, height = 1500 )
ggplot(crcdata_new, aes(fill=ghodsvali, x=year)) + 
    geom_bar(position="stack", stat="count")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + scale_x_continuous(breaks = crcdata_new$year) + scale_fill_manual(values = my_colors, name = "Ghodsvali Scale")
dev.off()

```

\newpage

## Stakeholder engagement by year

```{r echo=FALSE, warning=FALSE, message=FALSE}
#year
library(dplyr)
crcdata_st <- read.csv("./data/crc_data_ES2.csv")



my_colors <- c("#99CC99",  "#CC9966")

crcdata_stremove2011 <- subset(crcdata_st, year != "2011")
colnames(crcdata_stremove2011)[22] <- "Stakeholder Engagement"
ggplot(crcdata_stremove2011, aes(fill=`Stakeholder Engagement`, x=year)) + 
    geom_bar(position="stack", stat="count")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + scale_x_continuous(breaks = crcdata_st$year) + guides(col=guide_legend("Stakeholder Engagement? Yes or No")) + scale_fill_manual(values = my_colors)



```

\newpage

## Researcher types

```{r echo=FALSE, warning=FALSE, message=FALSE}
#researcher
df <- t(t(colSums(crcdata[,12:17])))
df <- df[order(df[,1], decreasing = TRUE),]
par(mar = c(6,4,4,4))
#barplot(df, las=2, main = "Researcher Types")

df <- data.frame(df)
rownames(df) <- c("Engineering", "Physics", "Imterdisciplinary", "Computer Science", "Math", "NGO")
df$group <- row.names(df)
df$group <- factor(df$group, levels=c("Engineering", "Physics", "Imterdisciplinary", "Computer Science", "Math", "NGO"))

ggplot(df, aes(x=group, y=df)) +
    geom_bar(stat='identity', fill="#99CC99")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + guides(col=guide_legend("Researcher Types")) +labs(y= "Count", x = "Researcher Type")


```
\newpage

## Stakeholder types

```{r echo=FALSE, warning=FALSE, message=FALSE}

#stakeholder types
df <- t(t(colSums(crcdata[,23:32])))
df <- df[order(df[,1], decreasing = TRUE),]
par(mar = c(11,4,4,4))
#barplot(df, las=2, main = "Stakeholder Types")


df <- data.frame(df)
rownames(df) <- c("Combined Government", "Combined Industry", "Farmers", "University", "Combined Coalition", "Public", "Experts", "Tribal Nations", "Youth", "Migrants")
df$group <- row.names(df)
df$group <- factor(df$group, levels=c("Combined Government", "Combined Industry", "Farmers", "University", "Combined Coalition", "Public", "Experts", "Tribal Nations", "Youth", "Migrants"))

ggplot(df, aes(x=group, y=df)) +
    geom_bar(stat='identity', fill="#99CC99")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + guides(col=guide_legend("Stakeholder Types")) +labs(y= "Count", x = "Stakeholder Type")

```
\newpage

## Ghodsvali scale breakdown

```{r echo=FALSE, warning=FALSE, message=FALSE}

#ghodsvali scale
df <- t(t(colSums(crcdata[,33:36])))
df <- df[order(df[,1]),]
par(mar = c(11,4,4,4))

#barplot(c(df[3], df[4], df[2], df[1]), las=2, main="Ghodsvali Scale Breakdown")

df <- data.frame(df)
rownames(df) <- c("Transformative", "Representation", "Nominal", "Instrumental")
df$group <- row.names(df)
library(dplyr)

# create a vector with letters in the desired order
x <- c("Nominal", "Instrumental",        "Representation",   "Transformative" )

df <- df %>%
  slice(match(x, group))



#rownames(df) <- c("Transformative", "Representation", "Nominal", "Instrumental")
#df$group <- row.names(df)
#df <- c(df[3,], df[4,], df[2,], df[1,])
df$group <- factor(df$group, levels=c("Nominal", "Instrumental", "Representation", "Transformative"))

ggplot(df, aes(x=group, y=df)) +
    geom_bar(stat='identity', fill="#99CC99")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + guides(col=guide_legend("Ghodsvali Scale Breakdown")) +labs(y= "Count", x = "Ghodsvali Scale")

```
\newpage

## Geographic location breakdown

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geography
df <- t(t(colSums(crcdata[,38:42])))
df <- df[order(df[,1], decreasing = TRUE),]
par(mar = c(8,4,4,4))
#barplot(df, las=2, main="Geographic Location Breakdown")




df <- data.frame(df)
rownames(df) <- c("Local", "Regional", "National", "Multinational", "Global")
df$group <- row.names(df)
library(dplyr)

# create a vector with letters in the desired order
#x <- c("Researcher", "Data Gathering",        "Inform",   "Envision", "Collaborative", "Empowered" )

#df <- df %>%
#  slice(match(x, group))



#rownames(df) <- c("Transformative", "Representation", "Nominal", "Instrumental")
#df$group <- row.names(df)
#df <- c(df[3,], df[4,], df[2,], df[1,])
df$group <- factor(df$group, levels=c("Local", "Regional", "National", "Multinational", "Global"))

ggplot(df, aes(x=group, y=df)) +
    geom_bar(stat='identity', fill="#99CC99")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + guides(col=guide_legend("Geography")) +labs(y= "Count", x = "Geography")

```


```{r echo=FALSE, warning=FALSE, message=FALSE}

#Odds of stakeholder scale predicting whether a solution was proposed or not

#set.seed(101) 
#sample = sample.split(crcdata, SplitRatio = .75)
#train = subset(crcdata, sample == TRUE)
#test  = subset(crcdata, sample == FALSE)


ghodsvali.fit<-glm(solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative,binomial(link = "logit"), data = crcdata)
# IAP2.fit<-glm(solution_proposed_YN ~STE_IAP2_data_gathering+STE_IAP2_inform+STE_IAP2_consult+STE_IAP2_involve+STE_IAP2_collab+STE_IAP2_empower,binomial(link = "logit"), data = crcdata)
# local.fit<-glm(solution_proposed_YN ~SC_researcher+SC_datagathering+SC_inform+SC_perspectives+SC_plan+SC_identify+SC_envision+SC_implement,binomial(link = "logit"), data = crcdata)

```


```{r echo=FALSE, warning=FALSE, message=FALSE}
#--sequentially compares the smaller model with the next more 
#--complex model by adding one variable in each step. Each of those 
#--comparisons is done via a likelihood ratio test 

#anova_ghodsvali_logit <- anova(ghodsvali.fit, test="Chisq")
#anova_IAP2_logit <- anova(IAP2.fit, test="Chisq")
#anova_local_logit <- anova(local.fit, test="Chisq")

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

#prediction_accuracy <- predict(ghodsvali.fit, test, type="response", interval="confidence")
```

\newpage
# Ghodsvali Scale Modeling - solution proposed

## Ghodsvali scale regression

Ghodsvali scale regression testing on whether a solution was proposed or not

```{r echo=FALSE, warning=FALSE, message=FALSE}

summary(ghodsvali.fit)

```
\newpage

## Ghodsvali scale odds

Odds of Ghodsvali scale predicting whether a solution was proposed or not

```{r echo=FALSE, warning=FALSE, message=FALSE}

ghodsvali.fit<-glm(solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative,binomial(link = "logit"), data = crcdata)
logistic.display(ghodsvali.fit)


```

\newpage

## Ghodsvali Ensembled Decision Tree with Feature Importance 

Ghodsvali ensembled decision tree with solution proposed being the dependent variable

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales

#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~ STE_G_nominal + STE_G_instrumental + 
                         STE_G_representation + STE_G_transformative,
  data = crcdata,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )

```

\newpage

# Stakeholder Engagement Modeling - solution proposed

## REGRESSION: Does engaging stakeholders increase the likelihood that a solution will be proposed/implemented?

Here we use classical logistic regression using a binomial function to determine if engaging stakeholders (Y/N) increases the odds that a solution will be proposed.


```{r echo=FALSE, warning=FALSE, message=FALSE}

#engagement of stakeholders vs. solutions

stakeholder_vs_solutions <-glm(solution_proposed_YN ~S_stakeholder_engagement_YN,family=binomial,data=crcdata)

summary(stakeholder_vs_solutions)
#logistic.display(stakeholder_vs_solutions)

```


\newpage

## ODDS RATIOS: Does engaging stakeholders increase the likelihood that a solution will be proposed/implemented?

Odds of whether engaging stakeholders increases the likelihood that a solution will be proposed?

```{r echo=FALSE, warning=FALSE, message=FALSE}

#engagement of stakeholders vs. solutions

stakeholder_vs_solutions <-glm(solution_proposed_YN ~S_stakeholder_engagement_YN,family=binomial,data=crcdata)

#summary(stakeholder_vs_solutions)
logistic.display(stakeholder_vs_solutions)

```

\newpage

# Diversity of stakeholders vs solution

## REGRESSION: Does the diversity of stakeholders increase the likelihood that a solution will be proposed?

Regression testing of whether Diversity of stakeholders predicts if a solution was proposed (Y/N).  In order to represent diversity, we have used a simple ratio calcuation which sums the number of stakeholders involved divided by the total number of possible stakeholder options.  A ratio which is closer to 1 has a greater level of stakeholder diversity.

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of stakeholders vs solutions (using ratio)

crcdata$ST_ratio <- rowSums(crcdata[,23:31])/9

stratio_vs_solutions <-glm(solution_proposed_YN ~ ST_ratio,family=binomial,data=crcdata)

summary(stratio_vs_solutions)
#logistic.display(stratio_vs_solutions)

```

\newpage

## ODDS RATIOS: Does the diversity of stakeholders increase the likelihood that a solution will be proposed?

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of stakeholders vs solutions (using ratio)

logistic.display(stratio_vs_solutions)

```
\newpage

## REGRESSION: If diversity of stakeholders does not increase proposing/implementing solutions, which stakeholders are more associated with proposing/implementing solutions?

Regression testing for which stakeholders predict whether a solution was proposed

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of stakeholders vs solutions (using ratio)

stratio_vs_solutions <-glm(solution_proposed_YN ~ ST_farmers + ST_combined_gov  + ST_combined_coalition + ST_combined_industry + ST_public + ST_university + ST_experts,family=binomial,data=crcdata)

summary(stratio_vs_solutions)

#logistic.display(stratio_vs_solutions)

```

\newpage

## ODDS RATIOS: stakeholders vs solution

Odds whether a specific stakeholder predicts if a solution was proposed (Y/N). 

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of stakeholders vs solutions (using ratio)

stratio_vs_solutions <-glm(solution_proposed_YN ~ ST_farmers + ST_combined_gov  + ST_combined_coalition + ST_combined_industry  + ST_public + ST_university + ST_experts,family=binomial,data=crcdata)

#summary(stratio_vs_solutions)
logistic.display(stratio_vs_solutions)

```

\newpage

## DECISION TREE: Ensembed Decision Tree - stakeholders vs solution


```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales

#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~ST_farmers + ST_combined_gov  + ST_combined_coalition + ST_combined_industry + ST_migrants + ST_public + ST_university + ST_experts,
  data = crcdata,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )






#-----end new



# library(randomForest)
# library(caret)
# set.seed(2223)
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# rf <- randomForest(solution_proposed_YN ~ST_farmers + ST_combined_gov  + ST_combined_coalition + ST_combined_industry + ST_migrants + ST_public + ST_university + ST_experts, data=train, proximity=TRUE, importance=TRUE)
# print(rf)
# 
# p2 <- predict(rf, test)
# confusionMatrix(p2, test$solution_proposed_YN)
# 
# hist(treesize(rf),
#      main = "No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf,
#            main = "Top 10 - Variable Importance")
# importance(rf)
# 
# #--
# 
# # Required libraries
# library(randomForest)
# library(ggplot2)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()



```


\newpage

# Researcher Modeling - solution proposed

## REGRESSION: Does researcher type increase the likelihood that a solution will be proposed?

Regression of whether researcher type predicts if a solution was proposed (Y/N). 

```{r echo=FALSE, warning=FALSE, message=FALSE}

#researchers type vs solutions

researchertype_vs_solutions <-glm(solution_proposed_YN ~ R_ngo +  R_eng + R_math + R_compsci +R_phys + R_interdis + R_socsci + R_economics + R_ag + R_other,family=binomial,data=crcdata)

summary(researchertype_vs_solutions)
#logistic.display(rratio_vs_solutions)

```
\newpage

## ODDS RATIOS: Does researcher type increase the likelihood that a solution will be proposed?

Odds of whether researcher type predicts if a solution was proposed (Y/N). A ratio which is closer to 1 has a greater level of researcher diversity.
```{r echo=FALSE, warning=FALSE, message=FALSE}

#researchers type vs solutions (using ratio)

researchertype_vs_solutions <-glm(solution_proposed_YN ~ R_ngo +  R_eng + R_math + R_compsci +R_phys + R_interdis + R_socsci + R_economics + R_ag + R_other,family=binomial,data=crcdata)

logistic.display(researchertype_vs_solutions)

```

\newpage

## DECISON TREE: Researcher Type Ensembed Decision Tree - researcher type vs solution

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales


#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~R_ngo +  R_eng + R_math + R_compsci +R_phys + R_interdis + R_socsci + R_economics + R_ag + R_other,
  data = crcdata,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )






#-----end new

# library(randomForest)
# library(caret)
# set.seed(2223)
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# rf <- randomForest(solution_proposed_YN ~R_ngo +  R_eng + R_math + R_compsci +R_phys + R_interdis + R_socsci + R_economics + R_ag + R_other, data=train, proximity=TRUE, importance=TRUE) 
# print(rf)
# 
# p2 <- predict(rf, test)
# confusionMatrix(p2, test$solution_proposed_YN)
# 
# hist(treesize(rf),
#      main = "No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf,
#            main = "Top 10 - Variable Importance")
# randomForest::importance(rf)
# 
# #----new model on data where there is no solution proposed
# 
# crcdata_solution <- subset(crcdata, solution_proposed_YN == "Y")
# crcdata_nosolution <- subset(crcdata, solution_proposed_YN == "N")
# crcdata_stakeholders <- subset(crcdata, S_stakeholder_engagement_YN == "Y")
# crcdata_nostakeholders <- subset(crcdata, S_stakeholder_engagement_YN == "N")
# sample_crcdata_nosolution <- sample_n(crcdata_nosolution, 30)
# 
# solution_proposed_YN_combo <- rbind(crcdata_solution, sample_crcdata_nosolution)
# 
# rf2 <- randomForest(solution_proposed_YN ~R_ngo +  R_eng + R_math + R_compsci +R_phys + R_interdis + R_socsci + R_economics + R_ag + R_other, data=solution_proposed_YN_combo, proximity=TRUE, importance=TRUE) 
# 
# # rf3 <- randomForest(solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative, data=solution_proposed_YN_combo, proximity=TRUE, importance=TRUE) 
# 
# p2_nosolution <- predict(rf2, crcdata_nosolution)
# #confusionMatrix(p2_nosolution, test$solution_proposed_YN)
# 
# 
# hist(treesize(rf2),
#      main = "Balanced Model - No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf2,
#            main = "Balanced Model - Top 10 - Variable Importance")
# importance(rf2)
# 
# #--
# 
# # Required libraries
# library(randomForest)
# library(ggplot2)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf2, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()


```




\newpage

# Researcher Diversity Modeling - solution proposed

## REGRESSION: Does the diversity of researchers increases the likelihood that a solution will be proposed?

Regression of whether Diversity of researchers predicts if a solution was proposed (Y/N). In order to represent diversity, we have used a simple ratio calculation which sums the number of researcher types involved, divided by the total number of possible researcher options.  A ratio which is closer to 1 has a greater level of researcher diversity.

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of researchers vs solutions (using ratio)


crcdata$R_ratio <- rowSums(crcdata[,12:21])/10

rratio_vs_solutions <-glm(solution_proposed_YN ~ R_ratio,family=binomial,data=crcdata)

summary(rratio_vs_solutions)
#logistic.display(rratio_vs_solutions)

```


\newpage

## ODDS RATIOS: Does the diversity of researchers increases the likelihood that a solution will be proposed?

Odds of whether Diversity of researchers predicts if a solution was proposed (Y/N). In order to represent diversity, we have used a simple ratio calculation which sums the number of researcher types involved, divided by the total number of possible researcher options.  A ratio which is closer to 1 has a greater level of researcher diversity.

```{r echo=FALSE, warning=FALSE, message=FALSE}

#Diversity of researchers vs solutions (using ratio)


crcdata$R_ratio <- rowSums(crcdata[,12:21])/10

rratio_vs_solutions <-glm(solution_proposed_YN ~ R_ratio,family=binomial,data=crcdata)

#summary(rratio_vs_solutions)
logistic.display(rratio_vs_solutions)

```


\newpage

# Stakeholder Engagement Modeling - Ghodsvali

## Regression Testing - Stakeholder type vs level of engagement (Ghodsvali)

```{r echo=FALSE, warning=FALSE, message=FALSE}

#stakeholder type vs level of engagement ghodsvali

ST_engagement<-lm(cbind(ST_farmers,ST_combined_gov,ST_tribal,ST_combined_coalition,ST_combined_industry,ST_migrants,ST_youth,ST_public,ST_university,ST_experts)~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative,data=crcdata)

summary(ST_engagement)


```

\newpage

## Regression Testing - Stakeholder type vs solution 

```{r echo=FALSE, warning=FALSE, message=FALSE}

#stakeholder type vs solution

ST_solution <- glm(solution_proposed_YN~ST_farmers+ST_combined_gov+ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts, family=binomial, data=crcdata)

summary(ST_solution)

```

\newpage

# Geographic Location Modeling - solution proposed

## REGRESSION: Does the geographic location of the study increase the likelihood of proposed/implemented solutions?

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geographic area vs solution

geog_solution <- glm(solution_proposed_YN~G_local+G_regional+G_national+G_multinational+G_global, family=binomial, data=crcdata)
summary(geog_solution)
#logistic.display(geog_solution)


```

\newpage

## ODDS RATIOS: Does the geographic location of the study increase the likelihood of proposed/implemented solutions?

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geographic area vs solution

logistic.display(geog_solution)


```

\newpage

## DECISON TREE: Geographic area Ensembed Decision Tree - Geographic area vs solution

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales

#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~G_local+G_regional+G_national+G_multinational+G_global,
  data = crcdata,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )






#-----end new

# library(randomForest)
# library(caret)
# set.seed(2223)
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# 
# 
# rf <- randomForest(solution_proposed_YN ~G_local+G_regional+G_national+G_multinational+G_global, data=train, proximity=TRUE, importance=TRUE) 
# print(rf)
# 
# p2 <- predict(rf, test)
# confusionMatrix(p2, test$solution_proposed_YN)
# 
# hist(treesize(rf),
#      main = "No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf,
#            main = "Top 10 - Variable Importance")
# importance(rf)
# 
# #----new model on data where there is no solution proposed
# 
# crcdata_solution <- subset(crcdata, solution_proposed_YN == "Y")
# crcdata_nosolution <- subset(crcdata, solution_proposed_YN == "N")
# crcdata_stakeholders <- subset(crcdata, S_stakeholder_engagement_YN == "Y")
# crcdata_nostakeholders <- subset(crcdata, S_stakeholder_engagement_YN == "N")
# sample_crcdata_nosolution <- sample_n(crcdata_nosolution, 30)
# 
# solution_proposed_YN_combo <- rbind(crcdata_solution, sample_crcdata_nosolution)
# 
# rf2 <- randomForest(solution_proposed_YN ~G_local+G_regional+G_national+G_multinational+G_global, data=solution_proposed_YN_combo, proximity=TRUE, importance=TRUE) 
# 
# # rf3 <- randomForest(solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative, data=solution_proposed_YN_combo, proximity=TRUE, importance=TRUE) 
# 
# p2_nosolution <- predict(rf2, crcdata_nosolution)
# #confusionMatrix(p2_nosolution, test$solution_proposed_YN)
# 
# 
# hist(treesize(rf2),
#      main = "Balanced Model - No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf2,
#            main = "Balanced Model - Top 10 - Variable Importance")
# importance(rf2)
# 
# #--
# 
# # Required libraries
# library(randomForest)
# library(ggplot2)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf2, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()


```


\newpage

# Regional Location Modeling - solution proposed

## REGRESSION: Does the regional location of the study increase the likelihood of proposed/implemented solutions? 

Regions were grouped in: Europe/Asia, Middle East/Global - and Other.  Results of this bias-reduced logistic regression indicate that region was significantly associated with whether a solution was proposed, Chisquare = 7.28, p = .026. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(brglm2)

#regional area vs solution


crcdata$L_region <- as.factor(crcdata$L_region)

crcdata$L_region_3 <- with(crcdata, case_when(
  L_region %in% c("Europe", "Asia") ~ "EuropeAsia",
  L_region %in% c("Global", "Middle East") ~ "GlobalME",
  TRUE ~ "Other"
))
crcdata$L_region_3 <- factor(crcdata$L_region_3)
crcdata$L_region_3 <- relevel(crcdata$L_region_3, ref = "Other")

m_3 <- glm(solution_proposed_YN ~ L_region_3,
           family = binomial, data = crcdata, method = "brglmFit")
summary(m_3)

```

\newpage

## ODDS RATIOS: Does the regional location of the study increase the likelihood of proposed/implemented solutions? 

Compared to other regions, cases from Europe and Asia had 3.77 times higher odds of proposing a solution (95% CI [1.35, 10.48], p = .011). Global and Middle East cases also had higher odds (OR = 3.35, 95% CI [0.97, 11.57]), though this effect was marginal (p = .056).

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geographic area vs solution

exp(cbind(OR = coef(m_3), confint(m_3)))


```

\newpage

## DECISON TREE: Region area Ensembed Decision Tree - Region area vs solution

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales

#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

crcdata_OHE <- model.matrix(~ L_region - 1, data = crcdata)
crcdata_omit <- na.omit(crcdata)
crcdata_OHE2 <- data.frame(solution_proposed_YN = crcdata_omit$solution_proposed_YN, crcdata_OHE)


ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~.,
  data = crcdata_OHE2,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata_OHE2, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata_OHE2$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )






#-----end new

# library(randomForest)
# library(caret)
# set.seed(2223)
# 
# crcdata$solution_proposed_YN <- as.factor(crcdata$solution_proposed_YN)
# levels(crcdata$solution_proposed_YN) = c("N", "Y")
# 
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# crcdata$solution_implemented_YN <- as.factor(crcdata$solution_implemented_YN)
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# train <- na.omit(train)
# test <- na.omit(test)
# 
# # One-hot encode L_region for training
# X_train <- model.matrix(~ L_region - 1, data = train)
# 
# # One-hot encode L_region for testing
# X_test <- model.matrix(~ L_region - 1, data = test)
# 
# # Combine predictors with the target variable
# train <- data.frame(solution_proposed_YN = train$solution_proposed_YN, X_train)
# test  <- data.frame(solution_proposed_YN = test$solution_proposed_YN, X_test)
# 
# # train and test currently look like:
# # train <- data.frame(solution_proposed_YN = train$solution_proposed_YN, X_train)
# # test  <- data.frame(solution_proposed_YN = test$solution_proposed_YN,  X_test)
# 
# # 1) Add any missing dummy columns to test (fill with 0)
# miss_in_test <- setdiff(names(train), names(test))
# miss_in_test <- setdiff(miss_in_test, "solution_proposed_YN")
# for (nm in miss_in_test) test[[nm]] <- 0
# 
# # 2) Drop any extra columns in test that aren’t in train
# extra_in_test <- setdiff(names(test), names(train))
# if (length(extra_in_test)) test[extra_in_test] <- NULL
# 
# # 3) Reorder columns in test to match train (very important)
# test <- test[, names(train)]
# 
# # 4) Sanity checks
# stopifnot(identical(names(train), names(test)))
# stopifnot(identical(levels(train$solution_proposed_YN), levels(test$solution_proposed_YN)))
# 
# 
# rf <- randomForest(solution_proposed_YN ~., data=train, proximity=TRUE, importance=TRUE, ntree=1000,
#   classwt = c(N = 1, Y = 10) ) 
# 
# print(rf)
# 
# p2 <- predict(rf, test)
# 
# confusionMatrix(p2, test$solution_proposed_YN)
# 
# hist(treesize(rf),
#      main = "No. of Nodes for the Trees",
#      col = "green")
# 
# # Gini (most stable)
# varImpPlot(rf, type = 2, n.var = 1, main = "Top 10 - Variable Importance")
# 
# #varImpPlot(rf,main = "Top 10 - Variable Importance")
# importance(rf)
# 
# library(ggplot2)
# imp <- data.frame(Variable = rownames(importance(rf)), importance(rf))
# ggplot(imp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
#   geom_col(fill = "steelblue") +
#   coord_flip() +
#   labs(
#     title = "Random Forest Variable Importance",
#     y = "Mean Decrease in Accuracy",
#     x = "Region"
#   ) +
#   theme_minimal()
# 
# 
# # Required libraries
# library(randomForest)
# library(ggplot2)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()


```



\newpage

<!-- # Multivariate Stakeholder Engagement Modeling - geographic area -->

<!-- ## Regression Testing - stakeholder type vs geographic area - interactions and effects -->

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geographic area vs solution

# geog_stakeholdertype <- lm(cbind(ST_farmers,ST_combined_gov,ST_tribal,ST_combined_coalition,ST_combined_industry,ST_migrants,ST_youth,ST_public,ST_university,ST_experts)~G_local+G_regional+G_national+G_multinational+G_global, data=crcdata)
# 
# summary(geog_stakeholdertype)

```

\newpage

<!-- # Multivariate Geographic Modeling - Ghodsvali -->

<!-- ## Regression Testing - Geographic area vs engagement (Ghodsvali) - interactions and effects -->

```{r echo=FALSE, warning=FALSE, message=FALSE}

#geographic area vs solution

# geog_engagement_ghodsvali <- lm(cbind(G_local,G_regional,G_national,G_multinational,G_global)~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative, data=crcdata)
# 
# summary(geog_engagement_ghodsvali)

```
\newpage

# DECISION TREE ANALYSIS - ALL VARIABLES 

## Looking at Decision Tree for all variables, including the Ghodsvali scale - with solution proposed as dependent variable

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales


#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

# 
# rf_weighted <- train(
#   solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative+ST_combined_gov+           ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio+R_ratio,
#   data = crcdata,
#   method = "ranger",
#   trControl = ctrl, 
#   tuneGrid = expand.grid(
#     mtry = 2,
#     splitrule = "gini",
#     min.node.size = c(20,30,40)
#   ),
#   num.trees = 1000,
#   importance = "impurity",
#   metric = "ROC"
# )



rf_weighted <- train(
  solution_proposed_YN ~R_ngo+R_eng+R_math+R_compsci+R_phys+R_interdis+R_socsci+R_economics+R_ag+R_other+STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative+ST_combined_gov+           ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio+R_ratio,
  data = crcdata,
  method = "ranger",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 20)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )






#-----end new

# library(randomForest)
# library(caret)
# set.seed(2223)
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# rf <- randomForest(solution_proposed_YN ~STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative+ST_combined_gov+           ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio, data=train, proximity=TRUE, importance=TRUE) 
# print(rf)
# 
# 
# #----new model on data where there is no solution proposed
# 
# hist(treesize(rf),
#      main = "Balanced Model - No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf,
#            main = "Balanced Model - Top 10 - Variable Importance")
# importance(rf)
# 
# #--
# 
# # Required libraries
# library(randomForest)
# library(ggplot2)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()
# 

```

\newpage

# DECISION TREE ANALYSIS - ALL VARIABLES - minus scaling

## Looking at Decision Tree for all variables - minus the Ghodsvali scale - with solution proposed as dependent variable

```{r echo=FALSE, warning=FALSE, message=FALSE}

#RF scales



#library(randomForest)
library(caret)
library(pROC)
#library(ROCR)
library(ranger)
library(ggplot2)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)

rf_weighted <- train(
  solution_proposed_YN ~R_ngo+R_eng+R_math+R_compsci+R_phys+R_interdis+R_socsci+R_economics+R_ag+R_other+ST_combined_gov+ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio+R_ratio,
  data = crcdata,
  method = "ranger",
  trControl = ctrl, 
  tuneGrid = expand.grid(
    mtry = 2,
    splitrule = "gini",
    min.node.size = c(20,30,40)
  ),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)

#model results for printout:

# ---- compact summary (one paragraph) ----
best <- rf_weighted$bestTune
res  <- rf_weighted$results
for (nm in names(best)) res <- res[res[[nm]] == best[[nm]], ]

auc_mean <- res$ROC[1]; auc_sd <- res$ROCSD[1]
trees <- rf_weighted$finalModel$num.trees
n <- nrow(rf_weighted$trainingData)
cls <- table(rf_weighted$trainingData$.outcome)

vi <- caret::varImp(rf_weighted, scale = TRUE)$importance
vi$Feature <- rownames(vi)
top3 <- paste(head(vi[order(-vi$Overall), "Feature"], 3), collapse = ", ")

cat(sprintf("**Random forest (ranger).** %d trees; mtry=%d; min.node.size=%d; 5-fold CV.\n\n",
            trees, best$mtry, best$min.node.size))
cat(sprintf("**CV AUC:** %.3f (SD=%.3f).  **n:** %d.  **Class counts:** %s.\n\n",
            auc_mean, auc_sd, n,
            paste(sprintf("%s=%d", names(cls), as.integer(cls)), collapse=", ")))
cat(sprintf("**Top features:** %s.\n", top3))






## 1) In-sample AUC (on the training data)
rf_probs <- predict(rf_weighted, newdata = crcdata, type = "prob")[, "Y"]
roc_train <- pROC::roc(crcdata$solution_proposed_YN, rf_probs, levels = c("N","Y"))
#auc(roc_train)

## 2) Cross-validated AUC from caret's saved out-of-fold predictions
## (works if you trained with savePredictions="final" in trainControl)

stopifnot(!is.null(rf_weighted$pred))  # will error if you didn't save predictions

cv_preds <- rf_weighted$pred

## keep only rows for the selected hyperparameters
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv_preds <- cv_preds[cv_preds[[nm]] == bt[[nm]], ]

roc_cv <- pROC::roc(cv_preds$obs, cv_preds$Y, levels = c("N","Y"))  # 'Y' column = P(class "Y")
#auc(roc_cv)


# Best hyperparameters chosen by caret
#rf_weighted$bestTune
#rf_weighted$results[order(-rf_weighted$results$ROC), ][, c("mtry","min.node.size","ROC","ROCSD")]

# Out-of-fold predictions for the best hyperparameters only
cv <- rf_weighted$pred
bt <- rf_weighted$bestTune
for (nm in names(bt)) cv <- cv[cv[[nm]] == bt[[nm]], ]

# CV AUC (preferred estimate)
roc_cv <- pROC::roc(cv$obs, cv$Y, levels = c("N","Y"))
#auc(roc_cv)

#plot(roc_cv, print.auc = TRUE, main = "Cross-Validated ROC (caret, best tune)")

# Assume roc_cv and roc_train already exist
auc_train <- round(pROC::auc(roc_train), 3)
auc_cv <- round(pROC::auc(roc_cv), 3)

plot(roc_cv, col = "red", lwd = 2, lty = 2,
     main = "Training vs CV ROC Curves")
lines(roc_train, col = "blue", lwd = 2)
grid()

legend("bottomright",
       legend = c(
         paste("CV (out-of-fold) AUC =", auc_cv),
         paste("Training AUC =", auc_train)
       ),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bty = "n")


# 1) Get permutation importance from the fitted caret model
vi <- varImp(rf_weighted, scale = FALSE)     # scale=TRUE rescales to [0,100]
imp <- vi$importance
imp$Feature <- rownames(imp)

# 2) Keep the best-tuned model's importance and sort
# (For binary classification, caret provides an "Overall" column)
imp <- imp[order(-imp$Overall), ]
top_n <- head(imp, 10)

# 3) Pretty plot
ggplot(top_n, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(width = 0.7, fill = "blue") +
  coord_flip() +
  labs(
    title = "Feature Importance",
    x = "Feature",
    y = "Feature Importance"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = ggplot2::margin(5.5,12,5.5,5.5),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  )

#---new



# library(randomForest)
# library(caret)
# set.seed(2223)
# ind <- sample(2, nrow(crcdata), replace = TRUE, prob = c(0.7, 0.3))
# train <- crcdata[ind==1,]
# test <- crcdata[ind==2,]
# rf <- randomForest(solution_proposed_YN ~ST_combined_gov+           ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio, data=train, proximity=TRUE, importance=TRUE) 
# print(rf)
# 
# 
# #----new model on data where there is no solution proposed
# 
# hist(treesize(rf),
#      main = "Balanced Model - No. of Nodes for the Trees",
#      col = "green")
# varImpPlot(rf,
#            main = "Balanced Model - Top 10 - Variable Importance")
# importance(rf)
# 
# 
# # Extract variable importance
# var_imp <- importance(rf, type = 2)  # type = 2 is MeanDecreaseGini
# var_imp_df <- data.frame(Variable = rownames(var_imp),
#                          MeanDecreaseGini = var_imp[, "MeanDecreaseGini"])
# 
# # Sort and select top 10
# var_imp_df <- var_imp_df[order(-var_imp_df$MeanDecreaseGini), ]
# top_vars <- head(var_imp_df, 10)
# 
# # Plot
# ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseGini), y = MeanDecreaseGini)) +
#   geom_bar(stat = "identity", fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Top 10 Variable Importance - Random Forest Model",
#        x = "Variable", y = "Mean Decrease in Gini") +
#   theme_minimal()

```
\newpage

## Representative Decision Tree Plot - Balanced Model - Minus Scaling

```{r echo=FALSE, warning=FALSE, message=FALSE}


library(rpart)
library(rpart.plot)

# # Make sure your response is a factor for classification
# crcdata$solution_proposed_YN <- as.factor(crcdata$solution_proposed_YN)
# 
# # Fit a decision tree using rpart
# tree_model <- rpart(
#   solution_proposed_YN ~ ST_combined_gov + ST_tribal + ST_combined_coalition +
#     ST_combined_industry + ST_migrants + ST_youth + ST_public + ST_university +
#     ST_experts + G_local + G_regional + G_national + G_multinational +
#     G_global + year + L_region_3 + ST_ratio,
#   data = crcdata,
#   method = "class",  # use "anova" for regression
#   control = rpart.control(cp = 0, minsplit = 2, maxdepth = 5)
# )
# 
# # Plot the tree
# rpart.plot(tree_model)



library(rpart)
library(rpart.plot)

probs <- predict(rf_weighted, newdata = crcdata, type = "prob")
pos   <- if ("Y" %in% colnames(probs)) "Y" else colnames(probs)[2]
pY    <- probs[[pos]]

tree_model <- rpart(
  pY ~ R_ngo+R_eng+R_math+R_compsci+R_phys+R_interdis+R_socsci+R_economics+R_ag+R_other+ST_combined_gov + ST_tribal + ST_combined_coalition +
    ST_combined_industry + ST_migrants + ST_youth + ST_public + ST_university +
    ST_experts + G_local + G_regional + G_national + G_multinational +
    G_global + year + L_region_3 + ST_ratio+R_ratio,
  data   = crcdata,
  method = "anova",
  control = rpart.control(cp = 0.001, minbucket = 15, maxdepth = 4)
)

rpart.plot(tree_model, type = 2, extra = 1, under = TRUE,
           fallen.leaves = TRUE, faclen = 0, digits = 3, box.palette = "GnBu",
           main = "Representative Tree")

# Fidelity to forest probs (lower Brier is better)
mean((predict(tree_model, crcdata) - pY)^2)


```

\newpage

## Representative Decision Tree Plot - All variables

```{r echo=FALSE, warning=FALSE, message=FALSE}

library(rpart)
library(rpart.plot)

set.seed(2223)

crcdata$solution_proposed_YN <- factor(crcdata$solution_proposed_YN, levels = c("N", "Y"))

crcdata_renamed <- crcdata

colnames(crcdata_renamed) <- c("Year", "Solution Proposed", "Solution Implemented", "Tech", "Policy", "Institutional", "Social", "Economic", "Ecological", "Educational", "Was a model used?", "NGO Researchers", "Engineering Researchers", "Math Researchers", "Computer Science Researchers", "Physics Researchers", "Interdisiplinary Researchers", "Social Science Researchers", "Economics Researchers", "Agriculture Researchers", "Other Researchers", "Were Stakeholders Engaged?", "Farmer Stakeholders", "Combined Government Stakeholders", "Tribal Stakeholders", "Combined Coalition Stakeholders", "Combined Industry Stakeholders", "Migrant Stakeholders", "Youth Stakeholders", "Public Stakeholders", "University Stakeholders", "Expert Stakeholders", "Ghodsvali Nominal Engagment", "Ghodsvali Instrumental Engagement", "Ghodsvali Representation Engagement", "Ghodsvali Transformative Engagement", "Non-Descriptive Geography", "Local Geography", "Regional Geography", "National Geography", "Multinational Geography", "Global Geography", "Geography Not Available", "Study Nationality", "Study Region", "Study Region Draft1","Study Combined Region Draft2", "Stakeholder Diversity Ratio", "Researcher Stakeholder Ratio", "Study Combined Region")

ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)


library(caret)

# make sure the outcome is a 2-class factor and set the positive class first
# crcdata_renamed$`Solution Proposed` <- relevel(
#   factor(crcdata_renamed$`Solution Proposed`), ref = "Yes"  # <-- your positive class
# )

# List your predictors (leave OUT the outcome)
x <- c("Year","Tech","Policy","Institutional","Social",
       "Economic","Ecological","Educational","Was a model used?","NGO Researchers",
       "Engineering Researchers","Math Researchers","Computer Science Researchers",
       "Physics Researchers","Interdisiplinary Researchers","Social Science Researchers",
       "Economics Researchers","Agriculture Researchers","Other Researchers",
       "Were Stakeholders Engaged?","Farmer Stakeholders","Combined Government Stakeholders",
       "Tribal Stakeholders","Combined Coalition Stakeholders","Combined Industry Stakeholders",
       "Migrant Stakeholders","Youth Stakeholders","Public Stakeholders","University Stakeholders",
       "Expert Stakeholders","Ghodsvali Nominal Engagment","Ghodsvali Instrumental Engagement",
       "Ghodsvali Representation Engagement","Ghodsvali Transformative Engagement",
       "Non-Descriptive Geography","Local Geography","Regional Geography","National Geography",
       "Multinational Geography","Global Geography","Geography Not Available",
       "Stakeholder Diversity Ratio","Researcher Stakeholder Ratio","Study Combined Region")

# Build a properly quoted formula
f <- reformulate(sprintf("`%s`", x), response = "`Solution Proposed`")

library(dplyr)

crcdata_renamed <- crcdata_renamed %>%
  select(`Solution Proposed`,
         where(~ all(!is.na(.))))


# TrainControl for ROC
ctrl <- trainControl(method = "cv", number = 5,
                     classProbs = TRUE, summaryFunction = twoClassSummary)

rf_weighted <- train(
  f, data = crcdata_renamed,
  method = "ranger",
  trControl = ctrl,
  tuneGrid = expand.grid(mtry = 2, splitrule = "gini",
                         min.node.size = c(20, 30, 40)),
  num.trees = 1000,
  importance = "impurity",
  metric = "ROC"
)
# 
# 
# rf_weighted <- train(Solution Proposed ~ Year+Solution Proposed+Solution Implemented+Tech+Policy+Institutional+Social+Economic+Ecological+Educational+Was a model used?+NGO Researchers+Engineering Researchers+Math Researchers+Computer Science Researchers+Physics Researchers+Interdisiplinary Researchers+Social Science Researchers+Economics Researchers+Agriculture Researchers+Other Researchers+Were Stakeholders Engaged?+Farmer Stakeholders+Combined Government Stakeholders+Tribal Stakeholders+Combined Coalition Stakeholders+Combined Industry Stakeholders+Migrant Stakeholders+Youth Stakeholders+Public Stakeholders+University Stakeholders+Expert Stakeholders+Ghodsvali Nominal Engagment+Ghodsvali Instrumental Engagement+Ghodsvali Representation Engagement+Ghodsvali Transformative Engagement+Non-Descriptive Geography+Local Geography+Regional Geography+National Geography+Multinational Geography+Global Geography+Geography Not Available+Study Nationality+Stakeholder Diversity Ratio+Researcher Stakeholder Ratio+Study Combined Region,  
#   data = crcdata_renamed,
#   method = "ranger",
#   trControl = ctrl,
#   tuneGrid = expand.grid(
#     mtry = 2,
#     splitrule = "gini",
#     min.node.size = c(20,30,40)
#   ),
#   num.trees = 1000,
#   importance = "impurity",
#   metric = "ROC"
# )

# 
# 
# rf_weighted <- train(
#   solution_proposed_YN ~R_ngo+R_eng+R_math+R_compsci+R_phys+R_interdis+R_socsci+R_economics+R_ag+R_other+STE_G_nominal+STE_G_instrumental+STE_G_representation+STE_G_transformative+ST_combined_gov+ST_tribal+ST_combined_coalition+ST_combined_industry+ST_migrants+ST_youth+ST_public+ST_university+ST_experts+G_local+G_regional+G_national+G_multinational+G_global+year+L_region_3+ST_ratio+R_ratio,
#   data = crcdata,
#   method = "ranger",
#   trControl = ctrl,
#   tuneGrid = expand.grid(
#     mtry = 2,
#     splitrule = "gini",
#     min.node.size = c(20,30,40)
#   ),
#   num.trees = 1000,
#   importance = "impurity",
#   metric = "ROC"
# )


probs <- predict(rf_weighted, newdata = crcdata_renamed, type = "prob")
pos   <- if ("Y" %in% colnames(probs)) "Y" else colnames(probs)[2]
pY    <- probs[[pos]]


tree_model <- rpart(
  f,
  data   = crcdata_renamed,
  method = "anova",
  control = rpart.control(cp = 0.001, minbucket = 10, maxdepth = 5)
)

jpeg("tree_highres_2.jpg", width = 6000, height = 4000, res = 600)  # large canvas + 600 dpi

rpart.plot(tree_model,
           type = 2,
           extra = 1,
           under = TRUE,
           fallen.leaves = TRUE,
           faclen = 0,
           digits = 3,
           box.palette = "GnBu",
           main = "Representative Tree",
           cex = 1)   # adjust text scaling (try 0.8–1.2)
dev.off()


rpart.plot(tree_model, type = 2, extra = 1, under = TRUE,
           fallen.leaves = TRUE, faclen = 0, digits = 3, box.palette = "GnBu",
           main = "Representative Tree")

# Fidelity to forest probs (lower Brier is better)
#mean((predict(tree_model, crcdata) - pY)^2)

```





